\pkg{mlr3spatiotempcv}: Spatiotemporal Resampling Methods for \pkg{mlr3}

### 2021-03-28 paper/paper.tex
add more method content

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_nsp <- rsmp("repeated_cv", folds = 4, repeats = 2)
R> rr_nsp <- resample(
+   task = task, learner = learner,
+   resampling = resampling_nsp
+ )
R> 
R> rr_nsp$aggregate(measures = msr("classif.auc"))
\end{CodeInput}
\begin{CodeOutput}
classif.auc 
  0.7600664 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_sp <- rsmp("repeated_spcv_coords", folds = 4, repeats = 2)
R> rr_sp <- resample(
+   task = task, learner = learner,
+   resampling = resampling_sp
+ )
R> 
R> rr_sp$aggregate(measures = msr("classif.auc"))
\end{CodeInput}
\begin{CodeOutput}
classif.auc 
  0.6100402 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_buffer <- rsmp("spcv_buffer", theRange = 1000)
R> 
R> autoplot(resampling_buffer, size = 0.7, task = task, fold_id = 1, crs = 4326) *
+   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
+   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
\end{CodeInput}

@@

``Spatial Blocking'' aims to create homogeneous blocks across the study
area dividing observations into different zones of which each represents
a partition. The way how these folds are created can be manifold:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Supplying a numeric value in meters via argument \texttt{theRange}
  will create a quadratic block pattern across the study area following
  either a random (\texttt{selection\ =\ "random"}), systematic
  (\texttt{selection\ =\ "systematic"}) or checkerboard structure
  (\texttt{selection\ =\ "checkerboard"}).
\end{enumerate}

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_block_random <- rsmp("spcv_block", range = 500, folds = 5)
R> 
R> autoplot(resampling_block_random, size = 0.8, fold_id = 1,
+          task = task, crs = 4326, show_blocks = TRUE) *
+   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
+   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
\end{CodeInput}
\begin{CodeOutput}
Warning in st_point_on_surface.sfc(sf::st_zm(x)): st_point_on_surface may not
give correct results for longitude/latitude data
\end{CodeOutput}

@@

@@

@@

>>>>

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_nsp <- rsmp("repeated_cv", folds = 4, repeats = 2)$instantiate(task)
R> rr_nsp <- resample(
+   task = task, learner = learner,
+   resampling = resampling_nsp
+ )
R> 
R> rr_nsp$aggregate(measures = msr("classif.auc"))
\end{CodeInput}
\begin{CodeOutput}
classif.auc 
  0.7600664 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_sp <- rsmp("repeated_spcv_coords", folds = 4, repeats = 2)$instantiate(task)
R> rr_sp <- resample(
+   task = task, learner = learner,
+   resampling = resampling_sp
+ )
R> 
R> rr_sp$aggregate(measures = msr("classif.auc"))
\end{CodeInput}
\begin{CodeOutput}
classif.auc 
  0.6100402 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_buffer <- rsmp("spcv_buffer", theRange = 1000)$instantiate(task)
R> 
R> autoplot(resampling_buffer, size = 0.7, task = task, fold_id = 1, crs = 4326) *
+   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
+   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
\end{CodeInput}

@@

``Spatial Blocking'' (\pkg{blockCV} package) creates homogeneous blocks
across the study area splitting observations into different zones
\citep[\citet{wenger2012a}]{bahn2012a}. Then, a number of zones is
combined randomly following desired the number of folds. For example, in
a five-fold scenario, the method tries to combines zones to end up with
partitions containing around 20\% of all observations.

The way how the blocks are created can be manifold:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Supplying a numeric value in meters via argument \texttt{theRange}
  (named just \texttt{range} in \{mlr3spatiotempcv\}) will create a
  quadratic block pattern across the study area. The aggregation of
  zones can either be random (default)
  (\texttt{selection\ =\ "random"}), systematic
  (\texttt{selection\ =\ "systematic"}) or checkerboard structure
  (\texttt{selection\ =\ "checkerboard"}).
\end{enumerate}

\texttt{selection\ =\ "random"}

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_block_random <- rsmp("spcv_block", range = 1000, folds = 5)
R> 
R> autoplot(resampling_block_random,
+   size = 0.8, fold_id = 1, task = task, crs = 4326,
+   show_blocks = TRUE, show_labels = TRUE) *
+   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
+   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
\end{CodeInput}

@@

\texttt{selection\ =\ "systematic"}: Here, aggregation follows the
linear count from \$1:n\_\{folds\} from top to bottom.

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_block_systematic <- rsmp("spcv_block",
+   range = 1000, folds = 5,
+   selection = "systematic")
R> 
R> autoplot(resampling_block_systematic,
+   size = 0.8, fold_id = 1, task = task, crs = 4326,
+   show_blocks = TRUE, show_labels = TRUE) *
+   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
+   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
\end{CodeInput}

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1} \end{center}

\texttt{selection\ =\ "checkerboard"}: This option ignores paramter
\texttt{range} and always creates two partitions following a ``1,2,1,2''
pattern.

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_block_checkerboard <- rsmp("spcv_block",
+   range = 1000, folds = 5,
+   selection = "checkerboard")
R> 
R> autoplot(resampling_block_checkerboard,
+   size = 0.8, fold_id = 1, task = task, crs = 4326,
+   show_blocks = TRUE, show_labels = TRUE) *
+   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
+   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
\end{CodeInput}

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1} \end{center}

The last method from the \pkg{blockCV} package, ``environmental
blocking'', makes use of \emph{k-means} clustering \citep{hartigan1979a}
in a possibly multivariate space. The user can specify a variable via
argument \texttt{feature} which levels will be used to create clusters.
Hereby, k-means will use Euclidean distance for cluster creation. Hence,
the selected input variables should be of type numeric. To avoid a
potential bias introduced by features with high variance when selecting
multiple features, all features are standardized by default.

The following example clusters by feature ``distance to forest''.

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_env <- rsmp("spcv_env",
+   features = "distdeforest", folds = 5)
R> 
R> autoplot(resampling_env,
+   size = 0.8, fold_id = 1, task = task, crs = 4326,
+   show_blocks = TRUE, show_labels = TRUE) *
+   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
+   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
\end{CodeInput}

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1} \end{center}

It is also possible to cluster by multiple features:

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_env_multi <- rsmp("spcv_env",
+   features = c("distdeforest", "slope"), folds = 5)
R> 
R> autoplot(resampling_env_multi,
+   size = 0.8, fold_id = 1, task = task, crs = 4326,
+   show_blocks = TRUE, show_labels = TRUE) *
+   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
+   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
\end{CodeInput}

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1} \end{center}

@@

The ``Spatial CV'' method after \citep{sperrorest} and implemented in
package \pkg{sperrorest} is similar to method ``environmental blocking''
in that it also uses \emph{k-means} clustering for fold creation.
However, it does not make use of arbitrary features but uses the
coordinates of all observations to create clusters in the spatial
domain. This is what lead to the naming of \texttt{spcv\_coords} in
\pkg{mlr3spatiotempcv}.

\begin{CodeChunk}
\begin{CodeInput}
R> resampling_coords <- rsmp("spcv_coords", folds = 5)$instantiate(task)
R> 
R> autoplot(resampling_coords,
+   size = 0.8, fold_id = 1, task = task, crs = 4326) *
+   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
+   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
\end{CodeInput}

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1} \end{center}

@@

Leave-Location-and-Time-Out from packge \pkg{CAST} is a spatiotemporal
resampling method. It requires both a spatial and a temporal variable
which are used to to separate observations in a multi-dimensional space.

In this example the \texttt{cookfarm} dataset is used because it
contains a temporal variable (``Date'') which is not present in dataset
\texttt{ecuador}.

The visualization is done in a dynamic way via \pkg{plotly} to visualize
the both temporal and spatial dimensions.

\begin{CodeChunk}
\begin{CodeInput}
R> task_spt = tsk("cookfarm")
R> resampling_cstf <- rsmp("sptcv_cstf",
+   folds = 5,
+   time_var = "Date", space_var = "SOURCEID")
R> 
R> autoplot(resampling_cstf,
+   fold_id = 2, task = task_spt, crs = 4326, static_image = TRUE,
+   file = "plot.pdf") *
+   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
+   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
\end{CodeInput}
\begin{CodeOutput}
NULL
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}

\begin{center}\includegraphics{plot} \end{center}
